---
title: "Stat 324 Homework #1 Due Wednesday Sept 14th 11:59 pm"
author: "Tom Rosen"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Submit your homework to Canvas by the due date and time. Email your lecturer if you have extenuating circumstances and need to request an extension. 

*If an exercise asks you to use R, include a copy of the code and output. Please edit your code and output to be only the relevant portions.

*If a problem does not specify how to compute the answer, you many use any appropriate method. I may ask you to use R or use manually calculations on your exams, so practice accordingly.

*You must include an explanation and/or intermediate calculations for an exercise to be complete.

*Be sure to submit the HWK1 Autograde Quiz which will give you ~20 of your 40 accuracy points.

*50 points total: 40 points accuracy, and 10 points completion

## Basics of Statistics and Summarizing Data Graphically (I)

**Exercise 1**. A number of individuals are interested in the proportion of citizens within a county who will vote to use tax money to upgrade a professional baseball stadium in the upcoming vote. Consider the following methods:

The **Baseball Team Owner** surveyed 8,000 people attending one of the baseball games held in the stadium. Seventy eight percent (78%) of respondents said they supported the use of tax money to upgrade the stadium.
 
The **Pollster** generated 1,000 random numbers between 1-52,661 (number of county voters in last election) and surveyed the 1,000 citizens who corresponded to those numbers on the voting roll. Forty three percent (43\%) of respondents said they supported the use of tax money to upgrade the stadium.

> a. What is the population of interest? What is the parameter of interest? Will this parameter ever be calculated?

The population of interest is the 8,000 people surveyed a the game and the general people of that county.  The parameter of interest is a new stadium funded by tax money.

> b. What were the sample sizes used and statistics calculated from those samples? Are these simple random samples from the population of interest? 

The sample size used was 1,000 people from both the pollster and 8,000 team survey. Of the pollster voters 43% supported the new stadium,
while at the team survey 78% supported the new stadium.  The team survey isn't random since it's supporters of the team who visited a game.  The Pollster was random since it was people randomly selected from the county.

> c. The baseball team owner claims that the survey done at the baseball stadium will better predict the voting outcome because the sample size was much larger. What is your response?

This is not necessarily true, since the individuals attended a game; clearly they are interested in baseball and the team.  Thus, it is more likely individuals at the stadium would want to improve it and spend tax money on it.  By surveying individuals in the county you get a more fair view on the situation.

\vspace{1cm}
\newpage

**Exercise 2:** After manufacture, computer disks are tested for errors. The table below tabulates the number of errors detected on each of the 100 disks produced in a day. 

| Number of Defects | Number of Disks |
| - | - |
|0 | 42 |
|1 | 30 |
|2 | 16 |
|3 | 7 |
|4 | 5 |

> a. Describe the type of data that is being recorded about the sample of 100 disks, being as specific as possible.

The data being represented here is quantitative, discrete data.  This is because were given the number of disks and the number of errors, 
making it quantitative, while there can't be a half disk, or half defect, making the data discrete.


> b.Code for a frequency histogram showing the frequency for number of errors on the 100 disks is given below. 

>> bi. Knit the document and confirm that the histogram displays in the knitted file.

```{r, echo=TRUE, eval=TRUE}
error.data=c(rep(0,42), rep(1,30), rep(2,16), rep(3,7), rep(4, 5))
hist(error.data,  breaks=c(seq(from=-0.5, 4.5, by=1)),
     xlab="Defects", main="Number of Defects", 
     labels=TRUE, ylim=c(0,50))

```

>> bii. Describe what the rep() function does in this code chunk.

It combines the data on the x axis with it's associated y value, enabling us to create the bars in the histogram.

>> biii. Describe how this breaks command affects the histogram's appearance in this code chunk.

It sets the x axis break points for the histogram.

>> biv. Describe how setting ylim=c(0,30) instead of ylim=c(0,50) would change the histogram's appearance. Which value for ylim is perferable for clear communication of the data? 

It cuts off the bar associated with 0 defects and makes the bar associated with 1 defect seem much larger than the other bars.
This skews the perception towards making the interpreter think there are far less defects than there actually are.  The current
value for ylim is preferable since it lets the viewer see the whole picture without, allowing them to more fairly interpret the data.  

\vspace{1cm}
