---
title: "Stat 324 Homework #2 Due Wednesday Sept 21st 11:59 pm"
author: "SOLUTIONS"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Submit your homework to Canvas by the due date and time. Email your lecturer if you have extenuating circumstances and need to request an extension. 

*If an exercise asks you to use R, include a copy of the code and output. Please edit your code and output to be only the relevant portions.

*If a problem does not specify how to compute the answer, you many use any appropriate method. I may ask you to use R or use manually calculations on your exams, so practice accordingly.

*You must include an explanation and/or intermediate calculations for an exercise to be complete.

*Be sure to submit the HWK2 Autograde Quiz which will give you ~20 of your 40 accuracy points.

*50 points total: 40 points accuracy, and 10 points completion

## Basics of Statistics and Summarizing Data Numerically and Graphically (I)


**Exercise 1**. There are 12 numbers in a sample, and the mean is $\bar{x}=27$. The minimum of the sample is accidentally changed from 13.8 to 1.38.

> a. Is it possible to determine the direction in which (increase/decrease) the mean ($\bar{x}$)changes? Or how much the mean changes? If so, by how much does it change? If not, why not?

Yes it is possible, since the mean is the calculated by taking the sum of the values, and dividing by the number of values, we know
that decreasing one of the values from 13.8 to 1.38 will decrease the mean.  We can't figure out by how much however, as we don't have access to the values to calculate the difference.

> b. Is it possible to determine the direction in which the median changes? Or how much the median changes? If so, by how much does it change? If not, why not?

The median will not change; since the value changed is a minimum value, and the median is the middle point in the data, changing the
minimum will have no effect on the median.

> c. Is it possible to predict the direction in which the standard deviation changes? If so, does it get larger or smaller? If not, why not? Describe why it is difficult to predict by how much the standard deviation will change in this case.

We can predict that the standard deviation will decrease because we've added a farther outlying point in the minimum direction, 
however, it's difficult to predict by how much it will decrease since the calculation for standard deviation is so complex
and takes into account multiple different factors like the data points, and the IQR.

\vspace{1cm}


**Exercise 2:** Recall the computer disk error data given used in HWK 1. The table below tabulates the number of errors detected on each of the 100 disks produced in a day. 

| Number of Defects | Number of Disks |
| - | - |
|0 | 42 |
|1 | 30 |
|2 | 16 |
|3 | 7 |
|4 | 5 |

A frequency histogram showing the frequency for number of errors on the 100 disks is given below. 

```{r, echo=TRUE, eval=TRUE}
error.data=c(rep(0,42), rep(1,30), rep(2,16), rep(3,7), rep(4, 5))
hist(error.data, breaks=c(seq(from=-0.5, 4.5, by=1)), xlab="Defects", main="Number of Defects", labels=TRUE, ylim=c(0,60))
```


> a. What is the shape of the histogram for the number of defects observed in this sample? Why does that make sense in the context of the question?

The histogram presented is in the shape of a right skew.  That makes sense, as the disk developer doesn't intend for defects,
so we would expect there to be more disks with few defects, with there being progressively fewer disks with multiple errors, 
creating the right skew we see.

> b. Calculate the mean and median number of errors detected on the 100 disks 'by hand' and using the built-in R functions. How do the mean and median values compare and is that consistent with what we would guess based on the shape? [You can use the text such as $\bar{x}=\frac{value1}{value2}$ to help you show your work neatly].

Mean = (Number of disks)/(Number of defect bins) = 100 / 5 = 20 = mean(c(42, 30, 16, 7, 5))

Median = Middle number of the pack, since there are 5 numbers, the median is 16 = median(c(42, 30, 16, 7, 5))

The mean and median I calculated is equivalent to the one produced by r, and is consistent with the shape of the graph.

> c. Calculate the sample standard deviation ``by hand" and using the built in R function. Are the values consistent between the two methods? How would our calculation differ if instead we considered these 100 values the whole population?  

sd(c(42, 30, 16, 7, 5)) = 15.76388

> d. Construct a boxplot for the number of errors data using R with helpful labels.  Explain how the shape of the data identified in (a) can be seen from the boxplot.

```{r}
boxplot(c(42, 30, 16, 7, 5), horizontal = TRUE, main = "Defects found on Disks", xlab = "Number of Defects", ylab = "Disks")
```

> e.  Explain why the histogram is better able to show the discrete nature of the data than a boxplot. 

The histogram does a better and more obvious job of showing how many disks had an exact number of errors.  The box plot is far more confusing, and doesn't make it obvious that most disks didn't have defects; if anything the box plot makes it seem as though .

\vspace{1cm}

>**Exercise 3:** A company that manufactures toilets claims that its new presure-assited toilet reduces the average amount of water used by more than 0.5 gallons per flush when compared to its current model. The company selects 20 toilets of the current type and 19 of the New type and measures the amount of water used when each toilet is flushed once.  The number of gallons measured for each flush are recorded below. The measurements are also given in flush.csv.

Current Model: 1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56

New Model: 1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36


>a. Use R to create histograms to display the sample data from each model (any kind of histogram that you want since sample sizes are very similar). Have identical x and y axis scales so the two groups' values are more easily compared. Include useful titles.

```{r}
hist(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56), xlab="Amount in Gallons", main="Current Model", labels=TRUE, ylim=c(0,10))

hist(c(1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36), xlab="Amount in Gallons", main="New Model", labels=TRUE, ylim=c(0,10))

#breaks=c(seq(from=0.7, 2.2, by=0.1)),
```

>b. Compare the shape of the gallons flushed by the two toilet models.

The old model has a right skew, while the current model is more symmetrical.

>c. Compute the mean and median gallons flushed for the Current and New Model toilets using the built-in R function. Compare both measures of center within each group and comment on how that relationship corresponds to the datas' shapes. Also compare the measures of center across the two groups and comment on how that relationship is evident in the histograms.

current: mean(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56)) = 1.5765

new: mean(c(1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36)) 
= 1.065789

current: median(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56)) = 1.595

new: median(c(1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36)) 
= 1.03



>d. Compute (using built-in R function) and compare the sample standard deviation of gallons flushed by the current and new model toilets. Comment on how the relative size of these values can be identified from the histograms. 

Old: sd(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56))= 0.2456843

New: sd(c(1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36))= 0.2058941



>e. Use R to create side-by-side boxplots of the two sets in R so they are easily comparable. 

```{r}
old_set <- c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56)
new_set <- c(1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36)
boxplot(old_set, new_set, horizontal = TRUE, main = "Gallons per Flush, New vs Current Model", names = c("Current", "New"), xlab = "Number of Gallons", ylab = "Models")
```

>f. Explain why there are no values shown as a dot on the Current Model flush boxplot. To what values do the Current model flush boxplot whiskers extend? (Use R for your boxplot calculations and type=2 for quantiles)



>g. What would be the mean and median gallons flushed if we combined the two data sets into one large data set with 39 observations? Show how the mean can be calculated from the summary measures in part (c) along with the sample sizes and explain why the median of the combined set cannot be computed based on (c).

mean(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56, 1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36)) = 1.328

median(c(1.63, 1.25, 1.23, 1.49, 2.11, 1.48, 1.94, 1.72, 1.85, 1.54, 1.67, 1.76, 1.46, 1.32, 1.23, 1.67, 1.74, 1.63, 1.25, 1.56, 1.28, 1.19, 0.90, 1.24, 1.00, 0.80, 0.71, 1.03, 1.27, 1.14, 1.36, 0.91, 1.09, 1.36, 0.91, 0.91, 0.86, 0.93, 1.36)) = 1.28

You can calculate the mean by taking the mean of the current and new toilets.
1.5765 + 1.065789 / 2 = 1.321144.
The median can't be combined set can't be computed based on C since we can only compare the median of the sets independently, and combining them and finding the median won't give you the median of a set with significantly more data points.

\vspace{.5cm}

**Exercise 4** You are adding Badger-themed bedazzle to your striped overalls and are using both red and white beads. You are interested in how the size of the bag of beads you select your beads from changes the probability of outcomes of interest. Compute the probability for outcomes a and b using three different sampling strategies each time.

(Small Pop) drawing without replacement from a small population where the bag of beads contains 6 White beads and 4 Red beads.

(Large Pop) drawing without replacement from a large population where the bag of beads contains 600 White beads and 400 Red beads.

(Same Pop) drawing from a population where the bag of beads always contains 60% White and 40% Red beads.

Example: Consider choosing 3 beads. Calculate the probability of selecting no white beads. 

Small Pop: P([RRR])=$\frac{4}{10}*\frac{3}{9}*\frac{2}{8}=0.03333333$

Large Pop: P([RRR])=$\frac{400}{1000}*\frac{399}{999}*\frac{398}{998}=0.06371181$

Same Pop: P([RRR])=$0.40*.40*.40=0.064$

```{r}
(4/10)*(3/9)*(2/8)
(400/1000)*(399/999)*(398/998)
.4*.4*.4

```

>a. Consider choosing 3 beads. Calculate the probability of selecting exactly 1 white bead. 

**Small Pop: ** (6/10) * (4/9) * (3/8) = 0.1

**Large Pop: ** (600/1000) * (400/999) * (399/998) = 0.09604795

**Same Pop: ** 0.60 * 0.40 * 0.40 = 0.096



>b. Consider choosing 3 beads. Calculate the probability of selecting at least 1 white bead.

**Small Pop: ** 1 - 0.03333333 = 0.9666667

**Large Pop: ** 1 - 0.06371181 = 0.9362882

**Same Pop: ** 1 - 0.064 = 0.936




>c. Consider sampling without replacement. Does drawing from a population that is **small** or **large** relative to your sample size result in an probability that is closest to the probability when sampling with replacement? 

When drawing from a population that is large relative to the sample size, the difference between sampling with and without replacement will be minimal.

